animbook: Visualizing changes in performance measures and demographic affiliations using animation

### 2024-02-19 animbook-journal.Rmd
19/02/2024 check on the paper

  The challenge in producing an animation like that in the New York Times article animation is the transformation of the data and connection with elements of the plot that will be animated. The complexities include allowing the user to choose the number of categories, standardizing distributions, and allowing the user to input pre-computed categorical data. These considerations provide the objective for creating an R package that can generalize the animation to suitably apply to a wide range of data.

  The structure of this paper is as follows. The next section explains the animation in The New York Times article and why it is relevant to the problem of studying zombie companies. It is followed by sections describing the data format needed, considerations about making animations, and visual design. The software and its functions and usage are detailed in the next section. The next section illustrates the usage of the software to examine company performance, particularly zombie patterns, and the changing political allegiance of voters in Australia. The discussion section summarizes, outlines some limitations, and suggests some future directions.

@@

  In the visualization, each observation is initially classified into one group at the start and potentially transitions into either the same group or a different group. This dynamics visualization constructs questions on the broader use of this visualization to other types of data. The potential use of this visualization on accounting data is to convey a message, as reported by the @oecd_report, that the concept of zombie companies is not unique to Japan alone. It is also present in the United States, which has a faster metabolize rate (more new listings and exits) relative to Japan.

  The political data that exhibits the movement of voters switching party affiliations between elections can be a valuable insight into the behavior of the voters. This data could be extended to incorporate demographic information about the voters, providing analysts with a significant insight into voter behavior. This allows them to consolidate effective campaigns for their political party. This also applies to marketing data, where customers shift their product interest to the competitor, providing the marketing analysts with an understanding of both the company's products and the overall market.

@@

  @Hvitfeldt_2018 provides R (@r) code to generate an animation similar to that of the New York Times. Taking inspiration from this example, \CRANpkg{animbook} expands upon the visualization in numerous ways and packages the code so that it can be use for a variety of similar animations for studying flow and temporal change data.

  The New York Times visualization is also similar to what is called a Sankey plot. Sankey plots are flow diagrams that show the movement/change from one category level to another, with the width of each arrow representing the proportional changes between levels. There are several R packages available to generate versions of Sankey plots, including \CRANpkg{ggalluvial} and \CRANpkg{ggparallel}. The main difference between the Sankey plot visualization and the animation featured in the New York Times lies in their methods of representing the proportion of changes. Sankey plots utilize the width of connecting lines, while the New York Times animation employs the density and color of points.

@@

  One of the frequently encountered values in the dataset is NA, commonly known as missing values. In the context of accounting data, NA might signify that a company is either not yet listed or has been de-listed. In scenarios like voter data, NA may indicate that individuals did not provide the necessary information. Disregarding NA values could result in the loss of valuable information. Therefore, instead of completely removing NA entries, this information is categorized into its quantiles, ensuring that the missing data are factored into the visualization.

@@

  The \CRANpkg{gganimate} package is an extension from the \CRANpkg{ggplot2} package to include the description of an animation. It added new grammar classes to the plot object, allowing it to understand how the plot should change over time. The use of `transition_*()` functions allows it to achieve this by specifying how the data evolves and how it relates to itself across time. This includes `gifski_renderer()` from the \CRANpkg{gifski} package (@gifski) to save animation in GIF format or `av_renderer()` from the \CRANpkg{av} package (@av) to save it into a video file format.

@@

  In the animation, the use of proportional shaded areas serves to enhance the comprehension of proportion information, effectively displaying the distribution within each group. To represent sub-groups, points are colored to focus attention on differences between groups of observation.

@@

From Figure \@ref(fig:data-diagram), there is a need to map numerical value to a category. One way to handle this is by ranking the sales and grouping the rankings into quantiles. In some cases, this may not be the best option. When the observation is moved up by quantile, one is bound to move down. This issue can be resolved by using an alternative method, which is grouping values based on their absolute values. Users may also be interested in grouping the data based on different groups, for example, ranking within a specific country. This generalization leads to four different scaling methods for the numerical data.

@@

  These are only the initial steps in formatting the data into a category. Now that there is a method to transform the data from the raw into a categorized format, the next step is to modify it into an animated data structure. It is carried out by assigning the frame to each individual observation, ensuring that each ID does not contain repeat frame values. It lets the \CRANpkg{gganimate} or \CRANpkg{plotly} perceive where the observation would be on the plot at a given frame, as seen in Figure \@ref(fig:animated-diagram).

@@

Once the data is prepared. The next step is to create the `ggplot` object as a basis for the animation. There are three plots available in this package. Two of the plots could be used for the animation, and another plot is used as a static visualization. All of the plots have an internal function that converts the categorized data format into the animated structure for each plotting function.

@@

  The `wallaby_data()` function is responsible for performing data manipulation and formatting tasks on the original object. It includes creating additional data components for labeling and shading. This function also responds by interpolating the non-linear path. It performs this task by mapping the `sine()` function provided in this package to each observation using the `map()` function from the \CRANpkg{purrr} package (@purrr). The frame is recalculated using the same method mentioned in the data preprocessing section. The `proportional_shade()` function is called to generate the proportional shaded data.

@@

It is necessary to save the plot as a `ggplot` object before passing it to the final function, `anim_animate()`, to animate. This function will automatically detect which rendering method was specified in the previous steps and add the minimum requirement functions accordingly. By default, if the user specifies the rendering as `gganimate`, then it will add the `transition_time()` function from the \CRANpkg{gganimate} package. Otherwise, the `animation_opts()` function will be added from the \CRANpkg{plotly} package.

@@

The `cat_change` data in the package is used to illustrate how to use this package. It has five categories, A to E, over two-time points. Since the data is already in the categorized data structure, the `anim_prep_cat()` function is used. The output from this function is a class called `categorized`.

@@

To obtain an object for the animation function, the `wallaby_plot` is used. The `subset` and `relation` arguments are applied to refine the final appearance of the animation, Figure \@ref(fig:top-one).

@@

Two examples of usage are provided. The osiris example studies company movement between quartiles of sales performance over a 12-year period. The voting example studies party preference changes between 2016 and 2019 elections.

@@

The Osiris data presented in this section was acquired from Bureau van Dijk (@bvd), compiling information on a total of 122,318 companies. It was subsequently a subset for the 2021 report, consisting of 1000 companies with 94 variables. The dataset included in the package focuses on 790 companies, spanning from 2006 to 2018. The variables in this dataset include year, ID, country, sales, and japan.

  @oecd_report reported that companies in the United States have a faster metabolic rate relative to Japanese companies. This means that US companies tend to move up and down performance metrics more frequently than Japanese companies. This can be examined using the wallaby and kangaroo plots. The wallaby plot allows the study of entry and exit behavior. The kangaroo plot allows for studying the full movement of companies over time.

@@

  In Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:osiris-figure}','\\@ref(fig:osiris-gif)'))`, the focus is placed on companies that exited the market in 2018 that were in the market at some point since 2006. The grey regions show the proportion of companies in each of the categories in 2006 on the left that exited the market by 2018. (explain the difference in proportion and not listed companies). In 2006, the companies represented by points, colored by country, were in one of the quartiles or "Not listed" category. All the companies in the Top 25% were US companies. Any Japanese companies in the Top 25% in 2006 did not exit the market (Not listed). But actually, there were non-Japanese companies in the Top 25% in 2006. It is also interesting that the larger proportion of the companies in the Top 25% are getting de-listed, and the lower the quartiles, the less likely that the companies will exit the market.

@@

  The package provides the tools to create animations for simple temporal patterns with a few steps, but these visualization methods have some limitations. It is primarily for communication rather than exploration because one typically knows what the main patterns are. If there are many categories (quantiles), the display will be too crowded. Because the animation involves perceiving patterns in groups of moving points, it can be challenging with unbalanced sample sizes. This requires the reader to mentally calculate the fraction of points that are in motion.

>>>>

  The challenge in producing an animation similar to the New York Times article animation is the transformation of the data and connection with elements of the plot that will be animated. The complexities include allowing the user to choose the number of categories, standardizing distributions, and allowing the user to input pre-computed categorical data. These considerations provide the objective for creating an R package that can generalize the animation to suitably apply to a wide range of data.

  The structure of this paper is as follows. The subsequent section explains the animation in The New York Times article and why it is relevant to the problem of studying zombie companies. It is then followed by sections describing the data format needed, considerations about making animations, and visual design. The functions and usage of the software are detailed in the next section. Just prior to the final section, the paper illustrates the usage of the software to examine company performance, particularly zombie patterns, and the changing political allegiance of voters in Australia. The concluding discussion section summarizes, outlines some limitations, and suggests future directions for the software.

@@

  In the visualization, each observation is initially classified into one group at the start and potentially transitions into either the same group or a different group. This dynamic visualization constructs questions on the broader use of this visualization to other types of data. The potential use of this visualization on accounting data is to convey a message, as reported by the @oecd_report, that the concept of zombie companies is not unique to Japan alone. It is also present in the United States, which has a faster metabolization rate (more new listings and exits) relative to Japan.

  The political data that exhibits the movement of voters switching party affiliations between elections can be a valuable insight into the behavior of the voters. This data could be extended to incorporate demographic information about the voters, providing analysts with significant insight into voter behavior and allowing them to consolidate effective campaigns for their political party. It can also be applied to marketing data, where customers shift their product interest to the competitor, providing the marketing analysts with an understanding of both the company's products and the overall market.

@@

  @Hvitfeldt_2018 provides R (@r) code to generate an animation similar to that of the New York Times. Taking inspiration from this example, \CRANpkg{animbook} expands upon the visualization in numerous ways and packages the code so that it can be used for a variety of similar animations for studying flow and temporal change data.

  The New York Times visualization is similar to a Sankey plot. Sankey plots are flow diagrams that show the movement/change from one category level to another, with the width of each arrow representing the proportional changes between levels. There are several R packages available to generate versions of Sankey plots, including \CRANpkg{ggalluvial} and \CRANpkg{ggparallel}. The main difference between the Sankey plot visualization and the animation featured in the New York Times lies in their methods of representing the proportion of changes. Sankey plots utilize the width of connecting lines, while the New York Times animation employs the density and color of points.

@@

  One of the frequently encountered values in the dataset is NA, commonly known as a missing value. In the context of accounting data, NA might signify that a company is either not yet listed or has been de-listed. In scenarios like voter data, NA may indicate that individuals did not provide the necessary information. Disregarding NA values could result in the loss of valuable information. Therefore, instead of completely removing NA entries, this information is categorized into its quantiles, ensuring that the missing data are factored into the visualization.

@@

  The \CRANpkg{gganimate} package is an extension from the \CRANpkg{ggplot2} package to include the description of an animation. It added new grammar classes to the plot object, allowing it to understand how the plot should change over time. The use of `transition_*()` functions allows it to achieve this by specifying how the data evolves and how it relates to itself across time. The \CRANpkg{gganimate} includes `gifski_renderer()` from the \CRANpkg{gifski} package (@gifski) to save animation in GIF format or `av_renderer()` from the \CRANpkg{av} package (@av) to save it into a video file format.

@@

  The incorporation of proportional shaded areas in this animation package serves to enhance the comprehension of proportion information, effectively displaying the distribution within each group. To represent sub-groups, points are colored to focus attention on differences between groups of observation.

@@

From Figure \@ref(fig:data-diagram), there is a need to map numerical value to a category. One way to handle this is by ranking the sales and grouping the rankings into quantiles. In some cases, this may not be the best option. When the observation is moved up by quantile, one is bound to move down, making the amount of observation in the quantile remain the same. This issue can be resolved by using an alternative method, which is grouping values based on their absolute values. Users may also be interested in grouping the data based on different groups, for example, ranking within a specific country. This generalization leads to four different scaling methods for the numerical data.

@@

  These are only the initial steps in formatting the data into a category. Now that there is a method to transform the data from the raw into a categorized format, the next step is to modify it into an animated data structure. It is carried out by assigning a frame to each individual observation, ensuring that each ID does not contain repeated frame values. It lets the \CRANpkg{gganimate} or \CRANpkg{plotly} perceive where the observation would be on the plot at a given frame, as seen in Figure \@ref(fig:animated-diagram).

@@

Once the data is prepared. The next step is to create the `ggplot` object as a basis for the animation. There are three plots available in this package. Two of the plots could be used for the animation, and another plot is used as a static visualization. All of each plot has an associated internal function that converts the categorized data format into the animated structure.

@@

  The `wallaby_data()` function is responsible for performing data manipulation and formatting tasks on the original object. It includes creating additional data components for labeling and shading. This function is also responsible for interpolating the non-linear path. It performs this task by mapping the `sine()` function provided in this package to each observation using the `map()` function from the \CRANpkg{purrr} package (@purrr). The frame is recalculated using the same method mentioned in the data preprocessing section. The `proportional_shade()` function is called to generate the proportional shaded data.

@@

It is necessary to store the plot as a `ggplot` object before passing it to the final function, `anim_animate()`, to animate. This function will automatically detect which rendering method was specified in the previous steps and add the minimum requirement functions accordingly. By default, if the user specifies the rendering as `gganimate`, then it will add the `transition_time()` function from the \CRANpkg{gganimate} package. Otherwise, the `animation_opts()` function will be added from the \CRANpkg{plotly} package.

@@

The `cat_change` data in the package is used to illustrate how to use this package. It has five categories, A to E, over two time points. Since the data is already in the categorized data structure, the `anim_prep_cat()` function is used. The output from this function is a class called `categorized`.

@@

To obtain an object for the animation function, the `wallaby_plot` function is used. The `subset` and `relation` arguments are applied to refine the final appearance of the animation, Figure \@ref(fig:top-one).

@@

Two real-world examples of usage are provided. The osiris example studies company movement between quartiles of sales performance over a 12-year period. The voting example studies party preference changes between the 2016 and 2019 elections.

@@

The Osiris data presented in this section was acquired from Bureau van Dijk (@bvd), compiling information on a total of 122,318 companies. It was subsequently subsetted for the 2021 report, consisting of 1000 companies with 94 variables. The dataset included in the package focuses on 790 companies, spanning from 2006 to 2018. The variables in this dataset include year, ID, country, sales, and japan.

  @oecd_report reported that companies in the United States have a faster metabolic rate relative to Japanese companies. This means that US companies tend to move up and down in the performance metrics more frequently than Japanese companies. This can be examined using the wallaby and kangaroo plots. The wallaby plot allows the study of entry and exit behavior. The kangaroo plot allows for studying the full movement of companies over time.

@@

  In Figure `r knitr::asis_output(ifelse(knitr::is_latex_output(), '\\ref{fig:osiris-figure}','\\@ref(fig:osiris-gif)'))`, the focus is placed on companies that exited the market in 2018 that were in the market at some point since 2006. The grey regions show the proportion of companies in each of the categories in 2006 on the left that exited the market by 2018. (explain the difference in proportion and not listed companies). In 2006, the companies represented by points, colored by country, were in one of the quartiles or "Not listed" category. All the companies in the Top 25% were US companies. Any Japanese companies in the Top 25% in 2006 did not exit the market (Not listed). It is worth noting that there are actually none of the Japanese companies in the Top 25% in 2006. It is also interesting that the larger proportion of the companies in the Top 25% are getting de-listed, and the lower the quartiles, the less likely that the companies will exit the market.

@@

  The package provides the tools to create animations for simple temporal patterns with a few steps, but these visualization methods have some limitations. This package is to be used primarily for communication rather than exploration because the end-user typically knows what the main patterns are. If there are many categories (quantiles), the display will be too crowded. Because the animation involves perceiving patterns in groups of moving points, it can be challenging with unbalanced sample sizes. This requires the reader to mentally calculate the fraction of points that are in motion.
